---
title: "mlflow + R: A Case Study"
output: html_notebook
---

## Introduction
A brief walkthrough of using `mlflow` from R for building robust machine
learning pipelines. This is meant to be a starting point rather than a
comprehensive guide. Full details about mlflow can be found in the [mlflow
documentation](https://www.mlflow.org/docs/latest/index.html).

## Environment
This example makes use of the following R packages:
  + `mlflow`
  + `carrier`
  + `tidyverse`
  + `tidymodels`
  + `pins`
  + `fs`
  + `randomForest`
The `mlflow` R package depends on `mlflow`, which can be installed by running
`mlflow::install_mlflow()`.

## Data
The example and associated data come from Julia Silge's updated [Supervised ML Course](https://supervised-ml-course.netlify.app/).

---

# Cars

```{r setup}
# Packages ----
library(mlflow)
library(carrier)
library(tidyverse)
library(tidymodels)
library(fs)
library(pins)

# Plotting ----
theme_set(theme_bw())
```

## mlflow setup
Create an mlflow experiment
```{r}
mlflow_set_experiment("cars2018")
artifact_dir <- dir_create("mlflow-artifacts")
```

## Data
```{r data}
cars2018 <- read_csv(pin("https://raw.githubusercontent.com/juliasilge/supervised-ML-case-studies-course/master/data/cars2018.csv"))
```

```{r}
(p <- ggplot(cars2018, aes(x = MPG)) +
  geom_histogram(bins = 25) +
  labs(title = "Fuel Efficiency",
       x = "Fuel efficiency (mpg)",
       y = "Number of cars"))

ggsave("mpg-hist.png", p, path = artifact_dir)
```

## tidymodels
### Data split
```{r}
set.seed(1234)
cars_vars <- cars2018 %>% 
  select(-Model, -`Model Index`)

car_split <- cars_vars %>% 
  initial_split(prop = 0.8, strata = Aspiration)

car_train <- training(car_split)
car_test <- testing(car_split)

glimpse(car_train)
glimpse(car_test)
```

### Model fit
#### Linear Model
```{r}


lm_mod <- linear_reg() %>% 
  set_engine("lm")

fit_lm <- lm_mod %>% 
  fit(log(MPG) ~ ., data = car_train)

fit_lm
```

#### Random Forest
```{r}
rf_mod <- rand_forest() %>% 
  set_engine("randomForest") %>% 
  set_mode("regression")

fit_rf <- rf_mod %>% 
  fit(log(MPG) ~ ., data = car_train)

fit_rf
```

#### Model Evaluation
```{r}
train_results <- car_train %>%
    mutate(MPG = log(MPG)) %>%
    bind_cols(predict(fit_lm, car_train) %>%
                  rename(.pred_lm = .pred)) %>%
    bind_cols(predict(fit_rf, car_train) %>%
                  rename(.pred_rf = .pred))

# Evaluate the performance
metrics(train_results, truth = MPG, estimate = .pred_lm)
metrics(train_results, truth = MPG, estimate = .pred_rf)
```

```{r}
test_results <- car_test %>%
    mutate(MPG = log(MPG)) %>%
    bind_cols(predict(fit_lm, car_test) %>%
                  rename(.pred_lm = .pred)) %>%
    bind_cols(predict(fit_rf, car_test) %>%
                  rename(.pred_rf = .pred))

# Evaluate the performance
metrics(test_results, truth = MPG, estimate = .pred_lm)
metrics(test_results, truth = MPG, estimate = .pred_rf)
```


### Model fit w/ mlflow
Use `with(mlflow_start_run())` to record details to a specific mlflow run
#### Linear Model
```{r}
with(mlflow_start_run(), {
  
  lm_mod <- linear_reg() %>% 
    set_engine("lm")
  
  fit_lm <- lm_mod %>% 
    fit(log(MPG) ~ ., data = car_train)
  
  train_results <- car_train %>%
    mutate(MPG = log(MPG)) %>%
    bind_cols(predict(fit_lm, car_train)) 
  
  test_results <- car_test %>%
    mutate(MPG = log(MPG)) %>% 
    bind_cols(predict(fit_lm, car_test))
  
  train_metrics <- metrics(train_results, truth = MPG, estimate = .pred)
  test_metrics <- metrics(test_results, truth = MPG, estimate = .pred)
  
  mlflow_log_metric("train_rmse", train_metrics %>% 
                      filter(.metric == "rmse") %>% 
                      pull(.estimate))
  mlflow_log_metric("test_rmse", test_metrics %>% 
                      filter(.metric == "rmse") %>% 
                      pull(.estimate))
  
  mlflow_predictor <- carrier::crate(~stats::predict(model, .x), model = fit_lm$fit)
  mlflow_log_model(mlflow_predictor, artifact_path = "model")
})
```

```{r}
mlflow_ui()
```

```{r}
model <- mlflow_load_model("/Users/jamesblair/Documents/RStudio/repos/mlflow-intro/R/mlruns/1/8d7ff225ac9643b29ed93d173d4f1874/artifacts/model")

mlflow_predict(model, car_test)
```



#### Random Forest
```{r}
with(mlflow_start_run(), {
  rf_mod <- rand_forest() %>% 
    set_engine("randomForest") %>% 
    set_mode("regression")
  
  fit_rf <- rf_mod %>% 
    fit_xy(y = log(car_train$MPG), x = car_train %>% select(-MPG))
  
  train_results <- car_train %>%
    mutate(MPG = log(MPG)) %>%
    bind_cols(predict(fit_rf, car_train)) 
  
  test_results <- car_test %>%
    mutate(MPG = log(MPG)) %>% 
    bind_cols(predict(fit_rf, car_test))
  
  train_metrics <- metrics(train_results, truth = MPG, estimate = .pred)
  test_metrics <- metrics(test_results, truth = MPG, estimate = .pred)
  
  mlflow_log_metric("train_rmse", train_metrics %>% 
                      filter(.metric == "rmse") %>% 
                      pull(.estimate))
  mlflow_log_metric("test_rmse", test_metrics %>% 
                      filter(.metric == "rmse") %>% 
                      pull(.estimate))
  
  mlflow_predictor <- crate(~randomForest:::predict.randomForest(model), model = fit_rf$fit)
  mlflow_log_model(mlflow_predictor, artifact_path = "model")
})
```

View the mlflow UI
```{r}
mlflow_ui()
```

Fetch model from mlflow
```{r}
model <- mlflow_load_model("/Users/jamesblair/Documents/RStudio/repos/mlflow-intro/R/mlruns/1/50d616c535734fe7920dee2e583a136b/artifacts/model")

mlflow_predict(model, car_test)
```


